Stateless Model Checking with Data-Race Preemption Points

### 2015-10-26 pldi-16/di.tex
do some anonymizing

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	%\includegraphics[width=0.15\textwidth]{tree0.pdf}
	%\includegraphics[width=0.225\textwidth]{tree1.pdf}
	%\includegraphics[width=0.25\textwidth]{tree2.pdf}
	%\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()} or {\tt cond\_wait()}~\cite{landslide}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}
\section{Design}

@@

It is quite simple to mechanically recognize when {\tt x} and {\tt y} correspond to different abstract allocations despite colliding on address.
\landslide~implements this check by adding a generation counter to its heap state tracking:
each allocation is given a unique ID,
and when evaluating whether two heap accesses can race,
the IDs of their containing blocks must match.
However, when limited to a single execution, suppressing any data race matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug}:
Now, the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Hence, reordering lines 6-7 before line 3 will cause {\tt x} and {\tt x2} to race.

>>>>

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	%\includegraphics[width=0.15\textwidth]{tree0.pdf}
	%\includegraphics[width=0.225\textwidth]{tree1.pdf}
	%\includegraphics[width=0.25\textwidth]{tree2.pdf}
	%\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()} or {\tt cond\_wait()}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}
\section{Design}

@@

It is quite simple to mechanically recognize when {\tt x} and {\tt y} correspond to different abstract allocations despite colliding on address.
We implemented this check by adding a generation counter to \landslide's heap tracking:
each allocation is given a unique ID,
and when evaluating whether two heap accesses can race,
the IDs of their containing blocks must match.
However, when limited to a single execution, suppressing any data race matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug}:
Now, the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Hence, reordering lines 6-7 before line 3 will cause {\tt x} and {\tt x2} to race.

### 2015-10-26 pldi-16/evaluation.tex
buncha minor revisions
do some anonymizing

We also counted the number of free-re-malloc false positives that \landslide~suppressed.

@@

To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported free-re-malloc data races during the first interleaving as {\em deterministic}, as a single-pass analysis must,
rather than waiting until future interleavings to confirm them (as explained in \sect{\ref{sec:free-re-malloc}}).

@@

% TODO: get these numbers
{\bf Finer-grained PP subsets.}
\quicksand~was able to partially guarantee safety in {\large \bf 99999\%} of tests with large maximal state spaces.
However, in {\large \bf 1337} tests, no more than the minimal state space could be verified,
and in {\large \bf 42} tests, not even that much.
Larger state spaces often result from finer-grained locking,
which can indicate a more complicated concurrent algorithm requiring more rigorous verification than a program with a single global lock.
Hence these corner cases are important to consider for future work.
While this work used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
we envision future Iterative Deepening implementations could incorporate this method to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space

>>>>

We also counted the number of malloc-recycle false positives that \landslide~suppressed.

@@

To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported malloc-recycle data races during the first interleaving as {\em deterministic}, as a single-pass analysis must,
rather than waiting until future interleavings to confirm them (as explained in \sect{\ref{sec:recycle}}).

@@

% TODO: get these numbers
{\bf Finer-grained PP subsets.}
\quicksand~was able to partially guarantee safety in {\large \bf 99999\%} of tests with large maximal state spaces.
However, in {\large \bf 1337} tests, no more than the minimal state space could be verified,
and in {\large \bf 42} tests, not even that much.
Larger state spaces often result from finer-grained locking,
which can indicate a more complicated concurrent algorithm requiring more rigorous verification than a program with a single global lock.
%Hence these corner cases are important to consider for future work.
While this work used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
we envision future Iterative Deepening implementations could incorporate this method to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space

### 2015-10-26 pldi-16/paper.tex
write about free-re-malloc and add figure 1
write algorithm 1, GetBestJob
minor revisions
do some anonymizing

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

@@

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}

% TODO: Finalize these numbers.
\newcommand\numthrlibs{79}
\newcommand\numpintoses{{\Large \bf NUMPINTOSES}}
\newcommand\numstudence{{\Large \bf NUMSTUDENCE}} % total pintoses plus p2s

@@

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
We dramatically reduce false-positive data race reports,
%by verifying the absence of bugs in their associated state spaces
and find crashes caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
Compared to SSS-MC, we find bugs faster in smaller subset state spaces,
%when the maximal state space is too large to explore
and find new bugs with data-race PPs that the maximal state space would not expose at all.

Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
We have also posted the log files from all our experiments.
They can be found here:

\url{https://github.com/bblum/landslide} % TODO: put QS in its own repository

\url{https://github.com/bblum/landslide} % TODO: put final logs somewhere

@@

Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics, upon which \landslide~is built.
Thanks to Ji\v{r}\'{i} \v{S}im\v{s}a and the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources

>>>>

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
%\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
%\doublespacing

@@

% TODO: Don't forget to run e.g. s/\\quicksand /\\quicksand~/
\newcommand\landslide{\textsc{Landslide}}
\newcommand\quicksand{\textsc{Quicksand}}
\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}

% TODO: Finalize these numbers.
\newcommand\numthrlibs{79}
\newcommand\numpintoses{79}
\newcommand\numstudence{158} % total pintoses plus p2s

@@

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
We dramatically reduce false-positive data race reports,
%by verifying the absence of bugs in their associated state spaces
and find crashes caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
% TODO: Not true; Replace instead with a claim about partial verification.
Compared to SSS-MC, we find bugs faster in smaller subset state spaces,
%when the maximal state space is too large to explore
and find new bugs with data-race PPs that the maximal state space would not expose at all.

Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
We have also posted our evaluation's full data set and logs.
They can be found here:

% TODO: Make data set, at least, public.
TODO: Data set at least can be public.

{\em github links removed for double-blind review}
%\url{https://github.com/bblum/landslide} % POST-SUBMISSION TODO: put QS in its own repository
%
%\url{https://github.com/bblum/landslide} % POST-SUBMISSION TODO: put final logs somewhere
% POST-SUBMISSION TODO: Don't forget to anonymize the names of all your grupos.

@@

%Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Many thanks to Vaishaal Shankar, Haryadi Gunawi, and David A. Eckhardt for generously providing student implementations from Berkeley's, U. of Chicago's, and CMU's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics, upon which \landslide~is built.
Thanks to
% TODO: Post-double-blind-review
% Ji\v{r}\'{i} \v{S}im\v{s}a and
the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources
the U.S. Army Research Office under grant number W911NF0910273.
